Bullet Groups as a test of $\Lambda$CDM

### 2014-03-06 paper/ms.tex
new results subsection

The Bullet Cluster provided a new kind of observational evidence of
the existece of dark matter. Quantifying the displacement between dark
matter and the dominant baryonic component (hot X-ray emitting gas)
has been used to test the CDM paradigm itself by quantifying the
substructure velocity requiered to produce such displacement and also
by estimating the expected abundance of such events in the Universe.

@@

As a first approximation there are two quantities that are available
from observations of these Bullet-like systems. The projected distance
between two dominand dark matter clumps and the ratio of the
galaxies' luminosities associated to them. From the simulation point
of view this can be translated into the 2D projected values of
$||\vec{r}||$, its value relative to the virial radius $X_{\rm 2D}=
||\vec{r}||_{\rm 2D}/R_{\rm vir}$ and the ratio of the circular velocities
of the two clumps $V_{\rm c,sub}/V_{\rm c,host}$. In a higher degree
of detail, in order to gaint better insight we use
the sub-structure velocity as a fraction of the host's circular
velocity, $||\vec{r}||/V_{\rm c, host}$, as a mesure of the the strenght
of the interaction.  Finally, we also measure the geometry of each
interaction through the values for $\mu$.

@@

To construct this figure we co-add all the halos in the sample (left,
groups; right, clusters) at all redshifts. We stack the data because
we do not observe any strong time evolution, aditionally this allows
us to increase the signal in each bin. Overplotted there is an circle
with error bars that represents the observational estimates for the
system \bullg.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figures_eps/figure_3.eps}
\end{center}
\caption{Integrated probability distribution for the relative velocity of the
sub-halo with respect to its host. The left panel shows the results
in physical units while the right panel show the same values
normalized by the circular velocity of the host halo. The line coding
follows the same structure as Figure \ref{fig:displacements}}
\label{fig:velocities}
\end{figure}

@@

Figure \ref{fig:} presents the geometry of the bullet groups in our 
 the relative geometry of the collision using the
variables $\mu$ and $X_{\rm off}$.

@@

\section{Discussion}
\label{sec:discussion}
Strictly speaking our results apply to multimodal groups and the e
expected separation betwen dark matter clumps. [...]

Expected displacement between dark and baryonic components are better
described by systems where the collision already occurred.[...]
(comparison with the old bullet clusters paper?)

The expectations to quantify this effects observationally are
high. [...]

@@

In order to explore the distribution of displacements expected in the observational, we define a new parameter given by:\\

\begin{equation}
 \frac{v_{circ,sub}}{v_{circ,halo}}=0.5
\end{equation}

Figure \ref{newparameter} shows the scatter plot of the parameter
$\left(\frac{v_{circ,sub}}{v_{circ,halo}}\right)$ vs the  displacement
($d_{2d,(X,Y)}$) between the host halo and the substructure and for
different redshifts. This parameter is  consistent with the
observations. The red star simbol in the Figure \ref{newparameter} is
equal to 0.54 corresponding to  the fraction in velocity dispersion in
the line-of-sigth of the group SL2S SJ08544-0121
($\sigma_{host,halo}=341^{+43}_{-109}$ kms$^{-1}$ and
$\sigma_{substructure}=185^{+30}_{-62}$), reported by
\citet{2013A&A...552A..80M}. Based in the parameter
$\left(\frac{v_{circ,sub}}{v_{circ,halo}}\right)>0.5$, we estimate the
expected distribution of displacements and this is shown in the Figure
\ref{displacements} for the two groups classified in this work.

>>>>

The Bullet Cluster provided a new kind of observational evidence of
the existece of dark matter. Quantifying the displacement between dark
matter and the dominant baryonic component (hot X-ray emitting gas)
\citep{ForeroRomero2011} has been used to test the CDM paradigm itself
by quantifying the substructure velocity requiered to produce such
displacement \citep{Hayashi2006} to finally estimate the expected
abundance of such events in the Universe.

@@

As a first approximation there are two quantities that are available
from observations of these Bullet-like systems. The projected distance
between two dominand dark matter clumps and the ratio of the
galaxies' luminosities associated to them. From the simulation point
of view this can be translated into the 2D projected values of
$||\vec{r}||$, its value relative to the virial radius $D_{\rm off}=
||\vec{r}||_{\rm 2D}/R_{\rm vir}$ and the ratio of the circular velocities
of the two clumps $V_{\rm c,sub}/V_{\rm c,host}$. In a higher degree
of detail, in order to gaint better insight we use
the sub-structure velocity as a fraction of the host's circular
velocity, $||\vec{r}||/V_{\rm c, host}$, as a mesure of the the strenght
of the interaction.  Finally, we also measure the geometry of each
interaction through the values for $\mu$.

@@

To construct this figure we co-add all the halos in the sample (left,
groups; right, clusters) at all redshifts. We stack the data because
we do not observe any strong time evolution, aditionally this allows
us to increase the signal in each bin. Overplotted there is an circle
with error bars that represents the observational estimates for the
system \bullg using the fraction in velocity dispersion in
the line-of-sigth of the group SL2S SJ08544-0121
($\sigma_{host}=341^{+43}_{-109}$\kms and
$\sigma_{sub}=185^{+30}_{-62}$\kms) reported by
\citet{2013A&A...552A..80M} and the DM displacement inferred from the
data presented by \cite{Gastaldello}.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figures_eps/figure_3.eps}
\end{center}
\caption{Integrated probability distribution for the relative velocity of the
sub-halo with respect to its host. The left panel shows the results
in physical units while the right panel show the same values
normalized by the circular velocity of the host halo. The line coding
follows the same structure as Figure \ref{fig:displacement}}
\label{fig:velocities}
\end{figure}

@@

@@

Figure \ref{fig:} presents the geometry of the bullet groups using the 
variables $\mu$ and $D_{\rm off}$. The first evident feature is that
most of the configurations have $|mu|>0.9$ ($\theta\leq 30^{\circ}$),
meaning that most of the collisions can be described as a head-on while
only a minority with $|mu|<0.9$ have more tangential trajectories. For
the pairs on radial trajectories there are three regions of interest
in this plane that describe different stages in the collision,
assuming that the sub-halo merges or falls below the detection
threshold after the first pass through the center of the host halo.

The first region has $\mu\approx-1$ and $D_{\rm off}>0.6$, which
locates the systems where a head-on collision has just started. The
sub-halo is close to the boundary of the host halo and is
infalling. The second region has $\mu\approx 1$ and $D_{\rm off}<0.6$;
at this stage the collision continues after the first crossing of the
host's center, the low number of halos with radial infalling velocities
and displacements $D_{\rm off}>0.6$ sugesst that this the maximum
range of radii for the apogee.  The third region corresponds to
$\mu\approx-1$ and $D_{\rm off}<0.6$ which corresponds to the
secondary infall after apogee.

\subsection{Displacement between Dark Matter and Baryons}
\label{sec:baryonic_displacements}

\begin{figure}
\begin{center}
%\includegraphics[width=0.5\textwidth]{Figures_eps/figure_3.eps}
\end{center}
\caption{Integrated probability distribution for the estimated
  baryonic displacements in the group and cluster samples at all redshifts.}
\label{fig:baryonic_displacements}
\end{figure}

These different collision geometries will produce different results in
terms of the displacement between dark matter and baryons. Strictly
speaking, the results we have derived so far, apply to multimodal groups and their
expected separation betwen the two dominant dark matter clumps but it
should not be interpreted that all have a corresponding DM-baryon
displacement. The systems where the halo is starting to fall into the
host ($\mu\approx-1$, $D_{\rm off}>0.6$) should not present a
detectable DM-baryon displacement.

We clearly expect such displacement in the case where the
sub-structure has already passed through the center of the host
halo. In this section we attempt an estimation for the DM-baryon
displacement statistics. We work under the following
hypothesis. First, we consider that systems with $|\mu|<0.9$ have a
baryonic displacement, $d_{2D}^{\prime}$, equal to zero. Second, we
consider that all systems with infalling velocities $\mu<-0.9$ and
large displacements $D_{\rm off}>0.6$ also have baryonic displacements
equal to zero. Third, for all the other cases we estimate the
displacement between the baryons and the dominand DM peak by the
distances between the dominant peak and the center of mass of the main
halo, $d_{2D}^{\prime}=D_{\rm off} R_{\rm vir}$, where $X_{\rm off}$
is the offset computed for each host halo in the original catalog.

This simplified modeldoes not take into account that there is a
fraction of halos with $\mu<-0.9$ and $D_{\rm off}<0.6$ for which the
collision has not started and should have $d_{2D}^{\prime}=0$. A detailed
modeling of this fraction requires the study of the complete merger
tree of the halo and sub-halo, a study beyond the scope of this
Letter. Instead we cautio the reader that the derived fraction of
halos with a displacement $<d_{2D}^{\prime}$ can be considered as an
upper limit.

The results for the integrated distributions for $>d_{2D}^{\prime}$
are shown in Figure \ref{fig:baryonic_displacements}.[..]

@@
